---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rbw: Residual Balancing Weights for Marginal Structural Models

Residual balancing is a method of constructing weights for marginal structural models, which can be used to estimate marginal effects of time-varying treatments and controlled direct/mediator effects in causal mediation analysis. Compared with inverse probability-of-treatment weights (IPW), residual balancing weights tend to be more robust and more efficient, and are easier to use with continuous exposures. This package provides three main functions, `rbwPoint()`, `rbwPanel()` and `rbwMed()`, that produce residual balancing weights for analyzing point treatments, time-varying treatments, and causal mediation, respectively.

**Reference**

- Zhou, Xiang and Geoffrey T Wodtke. 2020. "[Residual Balancing: A Method of Constructing Weights for Marginal Structural Models](https://doi.org/10.1017/pan.2020.2)" Political Analysis.

## Installation

You can install the released version of rbw from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("rbw")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("xiangzhou09/rbw")
```

## Estimating the Average Effect of a Point Treatment

The `rbwPoint()` function constructs residual balancing weights for estimating the average effect of a point treatment. The following example illustrates its use by estimating the average effect of televised political advertisements (`treat`) on campaign contributions (`Cont`) among 16,265 zipcodes in the 2004 and 2008 US presidential elections. 

```{r, message = FALSE, warnings = FALSE}
library(rbw)
# install.packages("survey")
library(survey)

# residual balancing weights
rbwPoint_fit <- rbwPoint(treat, baseline_x = c(log_TotalPop, PercentOver65, log_Inc, PercentHispanic, PercentBlack, density, per_collegegrads, CanCommute), data = advertisement)

# attach residual balancing weights to data
advertisement$rbw_point <- rbwPoint_fit$weights

# fit marginal structural model
rbw_design <- svydesign(ids = ~ 1, weights = ~ rbw_point, data = advertisement)

# the outcome model includes the treatment, the square of the treatment,
# and state-level fixed effects (Fong, Hazlett, and Imai 2018)
msm_rbwPoint <- svyglm(Cont ~ treat + I(treat^2) + factor(StFIPS), design = rbw_design)
summary(msm_rbwPoint)
```


## Estimating Marginal Effects of Time-varying Treatments

The `rbwPanel()` function constructs residual balancing weights for estimating marginal effects of time-varying treatments. The following example illustrates its use by estimating the effect of negative campaign advertising (`d.gone.neg`) on election outcomes (`demprcnt`) for 113 Democratic candidates in US Senate and Gubernatorial elections. 

```{r, message = FALSE}

# models for time-varying confounders
m1 <- lm(dem.polls ~ (d.gone.neg.l1 + dem.polls.l1 + undother.l1) * factor(week), data = campaign_long)
m2 <- lm(undother ~ (d.gone.neg.l1 + dem.polls.l1 + undother.l1) * factor(week), data = campaign_long)
xmodels <- list(m1, m2)

# residual balancing weights
rbwPanel_fit <- rbwPanel(treatment = d.gone.neg, xmodels = xmodels, id = id, time = week, data = campaign_long)

# merge weights into wide-format data
campaign_wide2 <- merge(campaign_wide, rbwPanel_fit$weights, by = "id")

# fit a marginal structural model (adjusting for baseline confounders)
rbw_design <- svydesign(ids = ~ 1, weights = ~ rbw, data = campaign_wide2)
msm_rbw <- svyglm(demprcnt ~ cum_neg * deminc + camp.length + factor(year) + office, design = rbw_design)
summary(msm_rbw)
```

## Estimating Controlled Direct Effects (CDE)

In causal mediation analysis, the `rbwMed()` function can be used to construct residual balancing weights for estimating the controlled direct effect or the controlled mediator effect with a marginal structural model. The following example illustrates its use by estimating the controlled direct effect of shared democracy (`democ`) on public support for war (`strike`) at different levels of perceived morality of war (`immoral`) for a sample of respondents in a survey experiment.

```{r, message = FALSE}
# models for post-treatment confounders
m1 <- lm(threatc ~ ally + trade + h1 + i1 + p1 + e1 + r1 +
  male + white + age + ed4 + democ, data = peace)

m2 <- lm(cost ~ ally + trade + h1 + i1 + p1 + e1 + r1 +
  male + white + age + ed4 + democ, data = peace)

m3 <- lm(successc ~ ally + trade + h1 + i1 + p1 + e1 + r1 +
  male + white + age + ed4 + democ, data = peace)

# residual balancing weights
rbwMed_fit <- rbwMed(treatment = democ, mediator = immoral,
  zmodels = list(m1, m2, m3), interact = TRUE,
  baseline_x = c(ally, trade, h1, i1, p1, e1, r1, male, white, age, ed4),
  data = peace)

# attach residual balancing weights to data
peace$rbw_cde <- rbwMed_fit$weights

# fit marginal structural model
rbw_design <- svydesign(ids = ~ 1, weights = ~ rbw_cde, data = peace)
msm_rbwMed <- svyglm(strike ~ democ * immoral, design = rbw_design)
summary(msm_rbwMed)
```

