---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rbw: Residual Balancing Weights for Marginal Structural Models

Residual balancing is a method of constructing weights for marginal structural models, which can be used to estimate marginal effects of time-varying treatments and controlled direct/mediator effects in causal mediation analysis. Compared with inverse probability-of-treatment weights (IPW), residual balancing weights tend to be more robust and more efficient, and are easier to use with continuous exposures. This package provides two main functions, `rbwPanel()` and `rbwMed()`, that produce residual balancing weights for analyzing time-varying treatments and causal mediation, respectively.

**Reference**

- Zhou, Xiang and Geoffrey T Wodtke. 2020. "[Residual Balancing: A Method of Constructing Weights for Marginal Structural Models](https://doi.org/10.1017/pan.2020.2)" Political Analysis.

## Installation

You can install the released version of rbw from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("rbw")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("xiangzhou09/rbw")
```

## Estimating Marginal Effects of Time-varying Treatments

The `rbwPanel()` function constructs residual balancing weights for estimating marginal effects of time-varying treatments. The following example illustrates its use by estimating the effect of negative campaign advertising (`d.gone.neg`) on election outcomes (`demprcnt`) for 113 Democratic candidates in US Senate and Gubernatorial elections. 

```{r, message = FALSE}
library(rbw)
# install.packages("survey")
library(survey)

# models for time-varying confounders
m1 <- lm(dem.polls ~ (d.gone.neg.l1 + dem.polls.l1 + undother.l1) * factor(week), data = campaign_long)
m2 <- lm(undother ~ (d.gone.neg.l1 + dem.polls.l1 + undother.l1) * factor(week), data = campaign_long)
xmodels <- list(m1, m2)

# residual balancing weights
rbwPanel_fit <- rbwPanel(exposure = d.gone.neg, xmodels = xmodels, id = id, time = week, data = campaign_long)

# merge weights into wide-format data
campaign_wide2 <- merge(campaign_wide, rbwPanel_fit$weights, by = "id")

# fit a marginal structural model (adjusting for baseline confounders)
rbw_design <- svydesign(ids = ~ 1, weights = ~ rbw, data = campaign_wide2)
msm_rbw <- svyglm(demprcnt ~ cum_neg * deminc + camp.length + factor(year) + office, design = rbw_design)
summary(msm_rbw)
```

## Estimating Controlled Direct Effects (CDE)

In causal mediation analysis, the `rbwMed()` function can be used to construct residual balancing weights for estimating the controlled direct effect or the controlled mediator effect with a marginal structural model. The following example illustrates its use by estimating the controlled direct effect of college education (`college`) on depression at age 40 (`cesd40`) at different levels of socioeconomic status (`ses`) for a subsample of respondents in the National Longitudinal Survey of Youth, 1979.

```{r, message = FALSE}
# models for post-treatment confounders
m1 <- lm(cesd92 ~ female + race + momedu + parinc + afqt3 +
  educexp + college, data = education)
m2 <- lm(prmarr98 ~ female + race + momedu + parinc + afqt3 +
  educexp + college, data = education)
m3 <- lm(transitions98 ~ female + race + momedu + parinc + afqt3 +
  educexp + college, data = education)

# residual balancing weights
rbwMed_fit <- rbwMed(treatment = college, mediator = ses,
  zmodels = list(m1, m2, m3), baseline_x = female:educexp,
  interact = TRUE, base_weights = weights, data = education)

# attach residual balancing weights to data
education$rbw <- rbwMed_fit$weights

# fit marginal structural model
rbw_design <- svydesign(ids = ~ 1, weights = ~ rbw, data = education)
msm_rbw <- svyglm(cesd40 ~ college * ses, design = rbw_design)
summary(msm_rbw)
```

